---
title: "SQL for DataScience"
author: "Pankaj Shah"
date: "6/1/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, error=FALSE, warningFALSE}
library(dplyr)
```

# SQL for DS
```{r, message=FALSE}
# Lets read the csv file from the source
econ_df <- readr::read_csv('https://raw.githubusercontent.com/shahnp/Teaching_course_materials/master/web.csv')
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
copy_to(con, econ_df)
```

# Head
```{r}
# Always better to look the data even before we do anything on it.
head(econ_df)
```

```{r}
library(DataExplorer)
df_info <- function(x) {
  data  <- as.character(substitute(x))  # data frame name
  size <- format(object.size(x), units="Mb")  # size (Mb)
  plot_missing(data.frame(x)) # Vizualization of Missing Data.
  
  ##column information
  column.info <- data.frame( column        = names(sapply(x, class)),
                             class         = sapply(x, class),
                             unique.values = sapply(x, function(y) length(unique(y))),
                             missing.count = colSums(is.na(x)),
                             missing.pct   = round(colSums(is.na(x)) / nrow(x) * 100, 2)) %>% arrange(desc(unique.values))
                            
  row.names(column.info) <- 1:nrow(column.info)
  list(data.frame     = data.frame(name=data, size=size),
       dimensions     = data.frame(rows=nrow(x), columns=ncol(x)),
       column.details = column.info)
}
df_info(econ_df) # Info about datasets.
```

```{r, error= FALSE, message= FALSE, warning= FALSE}
library(kableExtra)
t(apply(econ_df, MARGIN = 2, function(x) range(x, na.rm=TRUE))) %>% kable()
```

# SQL SELECT
```{r}
# Lets get data about all the device used for purchase
library(DBI)
dbGetQuery(con, "SELECT device FROM econ_df") %>% table()
```

# 3 way table
```{r}
# Lets group them looking at the purchase 
dbGetQuery(con, "SELECT referrer, device, purchase  FROM econ_df") %>% table()
```

# * 
```{r}
# Lets select everything from the datasets and se what other things we could look at.
dbGetQuery(con, "SELECT * FROM econ_df LIMIT 5")
```
# * LIMIT 
```{r}
# If the page is too long we can limit that to 10.
dbGetQuery(con, "SELECT * FROM econ_df limit 10") # LIMIT and limit is same not case sensitive but reserved words are always better in capital word.
```

# DISTINCT 
```{r}
# Lets see the referrer datasets as we have choosen distint it will all return value to 1.
dbGetQuery(con, "SELECT distinct referrer FROM econ_df") %>% table()

# once we remove distinct we will get a table 
dbGetQuery(con, "SELECT referrer FROM econ_df") %>% table()
```

# SLICE

## > (Greater than)
```{r}
# We can slice and dice based on Duration look at 468 apperance
dbGetQuery(con, "SELECT * 
                 FROM econ_df 
                 WHERE duration >= 468 
                LIMIT 5")
```

## == (Equal to)
```{r}
# Other way of doing slicing and dicing would be to used == sign

dbGetQuery(con, "SELECT * 
                 FROM econ_df
                 WHERE device == 'mobile' 
                 LIMIT 5")
```

##  AND OR 
```{r}
#  OR | AND ==  > 

dbGetQuery(con, "SELECT * 
                 FROM econ_df 
                 WHERE n_visit > 3 AND duration > 100 AND country == 'Ireland' OR country == 'France' LIMIT 5")

```

# BETWEEN OR
```{r}
# Combination of AND & OR
dbGetQuery(con, "SELECT * 
                 FROM econ_df
                 WHERE (n_visit == 5 OR n_visit == 3)  
                 AND (device = 'Mobile' OR device = 'tablet') LIMIT 5") # case sensitive 'Mobile'

dbGetQuery(con, "SELECT * 
                 FROM econ_df
                 WHERE (n_visit == 5 OR n_visit == 3)  
                 AND (device = 'mobile' OR device = 'tablet') LIMIT 5") # mobile -> double check with the data table.

# Between Two Numbers 
dbGetQuery(con, "SELECT * 
                 FROM econ_df
                 WHERE n_visit BETWEEN 1 AND 3 AND device = 'mobile' LIMIT 5")
```

# WHERE IN 
```{r}
# Choice of the number, character etc.

dbGetQuery(con, "SELECT * 
                 FROM econ_df 
                 WHERE n_visit IN (2, 4, 6, 8, 10) LIMIT 5")

dbGetQuery(con, "SELECT *
                 FROM econ_df 
                 WHERE n_visit IN (2,4,6,8,10) AND duration > 300 AND 
                       country IN ('China', 'Japan', 'Colombia') LIMIT 5")
```

# NULL 
```{r}
# No Null value  as we have seen in the beginning.
dbGetQuery(con, "SELECT * 
                 FROM econ_df 
                 WHERE device IS NULL") # Zero 
```

# LIKE (%)
```{r}
# % represents rest part of the word.
dbGetQuery(con, "SELECT * 
                 FROM econ_df  
                 WHERE country LIKE 'P%' LIMIT 5") # Starting with P 

dbGetQuery(con, "SELECT * 
           FROM econ_df
           WHERE country LIKE '%A'  LIMIT 5") # Ending in A not case sensitive.

```

# SUM 
```{r}
# 1.SUM
dbGetQuery(con, "SELECT SUM(n_visit) FROM econ_df")

# 2.SUM WHERE ==
dbGetQuery(con, "SELECT SUM(n_visit) 
                FROM econ_df
                WHERE referrer == 'direct'")

## 3.SUM  WHERE IN 
dbGetQuery(con, "SELECT SUM(order_items)
                 FROM econ_df
                 WHERE device IN ('tablet', 'laptop')") 

## 4.SUM Count GROUP ORDER
dbGetQuery(con, "SELECT device,SUM(order_items), 
                 Count(*) AS individual_device_group_by
                 FROM econ_df
                 GROUP BY device 
                 ORDER by individual_device_group_by DESC")

## 5. Count GROUP ORDER : No SUM 
dbGetQuery(con, "SELECT device, 
                 count(*) AS visits_device_group_by 
                 FROM econ_df
                 GROUP BY device 
                 ORDER by visits_device_group_by DESC")

## 6. > 
dbGetQuery(con, "SELECT SUM(n_visit) 
                 FROM econ_df
                 WHERE n_visit > 5")
```

# AVERAGE
```{r}
# AVERAGE
dbGetQuery(con, "SELECT AVG(n_visit) FROM econ_df")

# WHERE  LIKE
dbGetQuery(con, "SELECT AVG(n_visit) 
                 FROM econ_df 
                 WHERE country LIKE 'P%'")

# WHERE == 
dbGetQuery(con, "SELECT AVG(n_visit) AS avg_of_all_the_mobile 
                 FROM econ_df 
                 WHERE device == 'mobile'")
```

# MAX MIN 
```{r}
# MAX
dbGetQuery(con, "SELECT MAX(n_visit) FROM econ_df")

# MAX from single column
dbGetQuery(con, "SELECT MAX(n_visit) 
                 FROM econ_df
                 WHERE device == 'tablet'")

# Define the column name: AS 
dbGetQuery(con, "SELECT MAX(n_visit) AS max_visit  
                 FROM econ_df")

# MAX GROUP_BY ORDER_BY
dbGetQuery(con, "SELECT device, MAX(duration) AS max_duration_of_all_device 
                 FROM econ_df 
                 GROUP BY device 
                 ORDER by max_duration_of_all_device DESC")

# MIN 
dbGetQuery(con, "SELECT MIN(n_visit) FROM econ_df")

# MIN WHERE 
dbGetQuery(con, "SELECT MIN(n_visit) 
                 FROM econ_df 
                 WHERE duration BETWEEN 600 AND 900")

# MIN AS
dbGetQuery(con, "SELECT MIN(duration) AS min_duration_of_all_time
                 FROM econ_df")

```

# ORDER BY 
```{r}
# ORDER alphabetically countrywise
dbGetQuery(con, "SELECT * 
                 FROM econ_df 
                 ORDER BY country LIMIT 5")

# DESC ORDER by n_visit
dbGetQuery(con, "SELECT * 
                 FROM econ_df 
                 ORDER BY duration DESC LIMIT 5")
```

# Facebook posts by Members of the U.S. Congress in 2017

**SQLite database** 



```{r}
# Library
library(DBI)
library("odbc")
```


```{r}
# create database: this will create a file in our hard drive

db <- dbConnect(RSQLite::SQLite(), "facebook-db.sqlite")

# reading the first file
congress <- read.csv("https://raw.githubusercontent.com/shahnp/data/master/congress-facebook-2017.csv",stringsAsFactors=F)

# dbWriteTable : add dataframe to our database adding first table:
dbWriteTable(db, "congress", congress, overwrite = TRUE) # once it is written it is already there.

# testing that it works:
dbListFields(db, "congress")
dbGetQuery(db, 'SELECT * FROM congress LIMIT 5')
```

However, the files are too big to open them all in memory.
Instead, we will open them one by one, and then append them to the table.

```{r, warning= FALSE, error= FALSE, message= FALSE}
fls <- list.files("~/Desktop/SQL-workshop-master/data/posts", full.names=TRUE)

for (f in fls){
  
  message(f)

  # read file into memory
  fb <- read.csv(f, stringsAsFactors=F)
  
  # adding to table in SQL database
  dbWriteTable(db, "posts", fb, append=TRUE)
  
}

# testing that it works
dbListFields(db, "posts")
dbGetQuery(db, 'SELECT * FROM posts LIMIT 5')

# what if we make a mistake and want to remove the table?
# dbRemoveTable(db, "posts")

# and we close the connection for now
dbDisconnect(db)
```

# Querying an SQL database:

```{r}
db <- dbConnect(RSQLite::SQLite(), "facebook-db.sqlite")
test <- dbGetQuery(db, 'SELECT * FROM congress LIMIT 5')
glimpse(test)
```

```{r }
# test if we can extract any data.
dbGetQuery(db, "SELECT * FROM congress LIMIT 5")
```

```{r}
# Querying :one column
dbGetQuery(db, "SELECT name FROM congress LIMIT 5") # select certain column from the database.
```

```{r}
# Select multiple columns
dbGetQuery(db, "SELECT name, party FROM congress LIMIT 5")
```
# 
Lets look at the post 

```{r}
dbGetQuery(db, "SELECT * FROM posts LIMIT 5")
```

# UPPER
```{r}
dbGetQuery(db, "SELECT UPPER(message) FROM posts LIMIT 5")
```

```{r}
# We have a lower link which we made Upper adding expressions.
dbGetQuery(db, "SELECT from_name, likes_count/comments_count, UPPER(type) FROM posts LIMIT 5")
```

```{r}
# Adding aliases to the new columns : AS 
dbGetQuery(db, "SELECT from_name, likes_count/comments_count AS lIKE_RATIO FROM posts LIMIT 5")
```

```{r}

# Best way to write the code in multiple level so that user can read with ease.

# You can modify how to show the column name 
dbGetQuery(db, "SELECT LOWER(from_name),
           likes_count/comments_count AS like_ratio 
           FROM posts 
           LIMIT 5")
```

# Distinct
```{r}
# Unique values 
dbGetQuery(db, "SELECT DISTINCT from_name 
           FROM posts 
           LIMIT 5")
```

```{r}
# selecting based on values of a column
dbGetQuery(db, "SELECT name, party 
           FROM congress
           WHERE party = 'Republican'
           LIMIT 5")
```

```{r}
# working with dates greater than
dbGetQuery(db, "SELECT from_name, type, date 
           FROM posts
           WHERE date > '2017-01-01'
           LIMIT 5")


# Between two dates 
dbGetQuery(db, "SELECT from_name, type, date 
           FROM posts
           WHERE date BETWEEN '2017-01-01' AND '2017-01-03'
           LIMIT 5")
```

```{r}
# AND operator
dbGetQuery(db, "SELECT from_name, type, date, likes_count 
           FROM posts
           WHERE date > '2017-06-01' AND type != 'photo'AND likes_count > 500 
           ORDER by likes_count DESC
           LIMIT 5")
```

```{r}
# OR operator
dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE comments_count>100 AND (type = 'photo' OR type = 'video')
           LIMIT 5")
```

```{r}
# IN

dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE type IN ('video', 'event')
           LIMIT 5")
```

# Matching conditions _ %:
```{r}
# _ matches exactly one character:
dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE date LIKE '2017-01-__'
           LIMIT 5")

# % matches any number of characters:
dbGetQuery(db, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE date LIKE '2017-03%'
           LIMIT 5")
```


```{r}
# SQLite does not have Regular Expressions, but we can get creative.
dbGetQuery(db, "SELECT from_name, message, date
           FROM posts
           WHERE message LIKE '%hungary%'
           LIMIT 5")
```

# Group_by
```{r}
dbGetQuery(db, 
  "SELECT from_name, COUNT(*) AS post_count
  FROM posts
  GROUP BY from_name
  LIMIT 3")
```

```{r}
# sort : type_count by ORDER:
dbGetQuery(db, 
  "SELECT type, COUNT(type) AS type_count
  FROM posts
  GROUP BY type
  ORDER BY type_count LIMIT 5")

# now in descending orders
dbGetQuery(db, 
  "SELECT type, COUNT(type) AS type_count
  FROM posts
  GROUP BY type
  ORDER BY type_count DESC LIMIT 5")
```


```{r}
# top 3 most popular post?
dbGetQuery(db, 
  "SELECT from_name, message, likes_count, datetime
  FROM posts
  ORDER BY likes_count DESC
  LIMIT 3")
```

```{r}
# You can also specify the column number instead of the name
dbGetQuery(db, 
  "SELECT from_name, message, likes_count, datetime
  FROM posts
  ORDER BY likes_count DESC
  LIMIT 2")
```

```{r}
# what was the post with the highest comment to like ratio?
# We subset only posts with 1000 likes or more to avoid outliers.
dbGetQuery(db,
  "SELECT from_name, message, likes_count, comments_count, date,   
      comments_count/likes_count AS comment_like_ratio
  FROM posts
  WHERE likes_count > 1000
  ORDER BY comment_like_ratio DESC
  LIMIT 5")
```

# Join 

```{r}
library(DBI)
db <- dbConnect(RSQLite::SQLite(), "facebook-db.sqlite")

dbGetQuery(db, 
  "SELECT posts.likes_count, congress.party, posts.date
    FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
    LIMIT 5")
```

# ON 
```{r}
dbGetQuery(db, 
  "SELECT posts.from_name, posts.message, posts.shares_count, congress.party 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  WHERE party = 'Democrat'
  ORDER BY shares_count DESC
  LIMIT 3")
```

```{r}
dbGetQuery(db, 
  "SELECT posts.from_name, posts.message, posts.shares_count, congress.party 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  WHERE party = 'Republican'
  ORDER BY shares_count DESC
  LIMIT 3")
```

# Grouping and Aggregating
```{r}
# COUNT * = total no. of rows

dbGetQuery(db, 'SELECT COUNT(*) FROM posts')
dbGetQuery(db, 'SELECT COUNT(*) FROM congress')
```


```{r}
dbGetQuery(db, 
  "SELECT congress.party, COUNT(*) AS total_posts
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")

dbGetQuery(db, 
  "SELECT congress.party, congress.gender, COUNT(*) AS total_posts
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party, congress.gender")
```


```{r}
dbGetQuery(db, 
  "SELECT congress.party, domain, COUNT(*) AS domain_count
  FROM posts JOIN Congress
    ON congress.screen_name = posts.screen_name
  WHERE congress.party = 'Democrat'
  GROUP BY domain
  ORDER BY domain_count DESC
  LIMIT 5")
```

```{r}
dbGetQuery(db, 
  "SELECT congress.party, domain, COUNT(*) AS domain_count
  FROM posts JOIN Congress
    ON congress.screen_name = posts.screen_name
  WHERE congress.party = 'Republican'
  GROUP BY domain
  ORDER BY domain_count DESC
  LIMIT 5")
```

```{r}
# Average # of posts by party
dbGetQuery(db, 
  "SELECT congress.party, AVG(posts.likes_count), COUNT(*) 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")

# DIFFERENT WAY:
dbGetQuery(db, 
  "SELECT congress.party, SUM(posts.likes_count)/COUNT(*) AS average 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")
```


```{r}
# most popular post by party
dbGetQuery(db, 
  "SELECT from_name, message, congress.party, MAX(posts.likes_count), COUNT(*) 
  FROM posts JOIN congress
    ON congress.screen_name = posts.screen_name
  GROUP BY congress.party")
```

```{r}
# number of posts by day of the month
dbGetQuery(db, 
  "SELECT SUBSTR(date, 9, 10) AS day_of_month, COUNT(*) as post_count
  FROM posts
  GROUP BY day_of_month")
```

```{r, error= FALSE, message= FALSE, warning= FALSE}
library(sqldf)

sqldf("SELECT count(*) from congress")
```

